---
title: "Map trigpoints data"
author: "Phil Mike Jones"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Map trigpoints data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This article shows you how to plot the trigpoints data on an interactive map.
First we need to load some packages, and the trigpoints data itself (don't forget to install `trigpoints` if you haven't already):

```{r packages}
# install.packages("trigpoints")

library("dplyr")
library("sf")
library("leaflet")
library("trigpoints")
```

There are `r nrow(trigpoints)` trigpoints in the full data set---far too many to plot interactively---so I've filtered on those within 40km of Sheffield (completely arbitrarily).

```{r sheffield}
# There are too many trigpoints to plot nationally, so plot those around Shef
sheffield =
  data.frame(
    easting  = 435100,
    northing = 387100,
    stringsAsFactors = FALSE
  ) %>%
  st_as_sf(coords = c("easting", "northing"), crs = 27700)

trigpoints =
  trigpoints %>% 
  filter(st_within(., st_buffer(sheffield, 40000), sparse = FALSE))
```

I also remove those that are destroyed, and also those that are not the iconic [pillar type](http://trigpointing.uk/wiki/Pillar#Hotine_Pillar).

```{r filter-trigpoints}
trigpoints =
  trigpoints %>%
  filter(destroyed == 0L) %>%
  filter(type == "PILLAR")
```

Next I transform the coordinates from OSGB36 (British National Grid) to Mercator, so that they can be plotted on top of a base map.

```{r transform-mercator}
trigpoints =
  trigpoints %>%
  st_transform(crs = 4326)
```

Now we can create the plot.
First we add the markers which feature a `popup` when you click a trig point.
The popup will provide the name of the pillar, and also it's height in metres.
Finally we add the base map tiles.

```{r map-single-call, out.width="100%"}
map =
  leaflet(trigpoints) %>%
  addMarkers(popup = ~ paste0(
    htmltools::htmlEscape(name),
    "<br />",
    "Height: ",
    htmltools::htmlEscape(height), " m")
  ) %>% 
  addTiles()

map
```
